---
title: Data analysis for "Essentially blocked - Can structural factors block an essentialist interpretation of a formal explanation?"
author: "Marianna Zhang (marianna.zhang@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
---

## Introduction

<!-- Abstract.-->


<!-- A description of the stimuli and procedures in this experiment.-->


This project's [repository](https://github.com/mariannazhang/structuralblocking) and preregistration can be found online. 


## Methods

### Power Analysis

Nadya's estimate: 30-50 per cell?

Muradoglu et al (Exp 2 on gender)
93 5-6yo
formal explanation -> all 5 DVs - 0, 1 or 0, 0.5, 1
> Although the main effect of condition was not significant, p = .11, the condition by age interaction was, p = .044. Six-year-olds, but not 5-year olds, showed higher essentialism for properties introduced via formal explanations (ps = .011 and .999, respectively; see Figure 2). Children also gave more essentialist responses with age, p = .003.

Vasilyeva et al
41 4-5yo, 48 5-6yo, 67 adults
structural context -> innateness_switch - 1, 2, 3, 4
> an ANOVA with framing condition and age group as between-subjects factors revealed the predicted main effect of framing, F(1, 85) = 8.95, p = .004, partial_eta_p_sq = .095, with no main effect of age group, F(1, 85) = 1.05, p = .309, nor interaction, F(1, 85) = .01, p = .984. Similarly, adults rated the target property as more mutable under the structural than nonstructural framing, t(65) = 8.04, p = .001, d = 2.00.

```{r}
# pwr.t.test(d = cohensD(essentialism ~ context * explanation, data = df.means_subj),
#            power = 0.8,
#            sig.level = 0.05)
```


### Planned Sample

<!-- Planned sample size and/or termination rule, sampling frame, known demographics if any, preselection rules if any.-->

Planned sample size is 5-year-olds and 6-year-olds recruited from a university preschool and museums, as well as adults recruited on Amazon Mechanical Turk. MTurk participants will be required to be located within the United States, and will be required to have a HIT acceptance rate of 80% or above.

### Materials



### Procedure	



The adult survey paradigm can be found on Qualtrics. 

### Analysis Plan

<!-- data cleaning rules, data exclusion rules, covariates, etc. Key analysis of interest. Can also pre-specify additional analyses you plan to do.-->

Essentialism measure for each participant calculated as average of all 5 essentialism measures.

#### Exclusion criteria
Experimentor error, out of age range. 

#### Analysis of interest
ANOVA on essentialism measure with an interaction term for context and explanation. 





### Methods Addendum (Post Data Collection)
#### Actual Sample
<!-- actual sample size, actual power, exclusions, demographics --> 



#### Differences from pre-data collection methods plan





## Results


### Data preparation

Data preparation as specified in the analysis plan.

```{r "knitr config", cache = FALSE}
require("knitr")
```


```{r}
#### Load relevant libraries and functions
library(tidyverse)


#### Import data for analysis
df.data <- read_csv("../data/FYP_data_pilot_children_1.csv")
```


```{r}
#### Data exclusion / filtering
# Record exclusions: experimentor error, age
df.excl <- tibble(
  exp_error = sum(df.data$exp_error == "yes"),
  age = sum(df.data$age < 5 | df.data$age >= 7)
  )

# Exclude subjects: experimentor error, age
df.data <- df.data %>%
  filter(exp_error == "No" & 
           (age >= 5 & age < 7)) %>%
  select(-exp_error, -starts_with("check")) # Delete exp_error, check columns
```


```{r}
#### Memory check analysis






```


```{r}
#### Demographics analysis
# Add age categorical variable
df.data <- df.data %>% 
  mutate(age_cat = case_when(
    age < 5 ~ "4",
    age >= 5 & age < 6 ~ "5",
    age >= 6 & age < 7 ~ "6",
    age >= 7 ~ "7"))
# Age
df.dem_age <- df.data %>%
  count(age_cat)

# Gender
df.dem_gender <- df.data %>%
  count(gender)
```



```{r}
#### Prepare data for analysis

# Recode essentialism dependent measures
df.data$stability_past <- df.data$stability_past %>% 
  recode("No" = 0,
         "Maybe" = 0.5,
         "Yes" = 1)
df.data$stability_future <- df.data$stability_future %>% 
  recode("No" = 0,
         "Maybe" = 0.5,
         "Yes" = 1)
df.data$innateness_stop <- df.data$innateness_stop %>% 
  recode("No" = 1,
         "Maybe" = 0.5,
         "Yes" = 0)
df.data$innateness_switch <- 
  if_else(df.data$cb == "girlsYellow", 
          (recode(df.data$innateness_switch,
                  "For sure Green-Ball" = 0,
                  "Maybe Green-Ball" = 1/3,
                  "Maybe Yellow-Ball" = 2/3,
                  "For sure Yellow-Ball" = 1)),
          (recode(df.data$innateness_switch,
                  "For sure Green-Ball" = 1,
                  "Maybe Green-Ball" = 2/3,
                  "Maybe Yellow-Ball" = 1/3,
                  "For sure Yellow-Ball" = 0)))
df.data$inductivePoten <- df.data$inductivePoten %>% 
  recode("Just this girl" = 0,
         "A few girls" = 0.5,
         "A whole lot of girls" = 1)

# Gather to tidy long form
df.tidy <- df.data %>% 
  select(-starts_with("check"), -ends_with("expl")) %>% 
  gather(question, response, "stability_past":"innateness_switch")

# Order questions
df.tidy$question <-  df.tidy$question %>% 
  factor(levels=c('stability_past','stability_future','innateness_stop','inductivePoten', 'innateness_switch'))

# Calculate overall essentialism measure per subject
df.means_subj <- df.tidy %>% 
  group_by(subject, age, context, explanation) %>% 
  summarize(essentialism = mean(response, na.rm = TRUE))
```

### Preliminary look at data and visualization
```{r}
# Overall essentialism measure across subjects by condition
df.means <- df.tidy %>%
  group_by(context, explanation) %>% 
  summarize(essentialism = mean(response, na.rm = TRUE),
            sd = sd(response, na.rm = TRUE), 
            n = length(unique(subject)))
df.means
```

```{r}
# Set visualization theme
theme_set(
  ggthemes::theme_few()
)
```

```{r}
# Essentialism by context and explanation
ggplot(df.means_subj, aes(x = context, y = essentialism, fill = explanation, color = explanation)) +
  geom_point(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
             alpha = 0.8) + 
  stat_summary(fun.data = "mean_cl_boot",
               position = position_dodge(width = 0.8),
               geom = "linerange",
               size = 1,
               alpha = 0.5) +
  stat_summary(fun.y = "mean",
               position = position_dodge(width = 0.8),
               geom = "point",
               shape = 21,
               color = "black",
               size = 4) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(limits = c(0, 1))
ggsave("FYP_pilot_children_1_contextExpl.png")

# Essentialism by explanation over age, faceted by context
ggplot(df.means_subj, aes(x = age, y = essentialism, fill = explanation, color = explanation)) +
  facet_wrap(~ context) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = FALSE) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(limits = c(0, 1))

# Essentialism by context and explanation, faceted by DV
ggplot(df.tidy, aes(x = context, y = response, fill = explanation, color = explanation)) +
  facet_wrap(~ question) +
  geom_point(alpha = 0.4,
             position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0.05, dodge.width = 0.4)) + 
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(limits = c(0, 1))
ggsave("FYP_pilot_children_1_contextExpl_DV.png")

# Essentialism by context and explanation, faceted by DV
ggplot(df.tidy, aes(x = response, fill = explanation)) +
  facet_grid(context ~question) +
  geom_histogram(alpha = 0.4) + 
  scale_fill_brewer(palette = "Set1") +
  scale_x_continuous(breaks = pretty_breaks())
ggsave("FYP_pilot_children_1_contextExpl_DV_hist.png")
```

### Target analysis

The analyses as specified in the analysis plan. 

```{r}
df.means_subj %>% 
  aov(essentialism ~ 1 + context * explanation, .) %>% 
  summary()
```


### Exploratory analyses
```{r}

```




## Discussion



